---
title: "ProJIVE and CJIVE"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# Short Introudction
This document serves as an illustrative example of JIVE analysis using both Canonical JIVE (i.e. CJIVE which is equivalent to AJIVE given the same joint and individual ranks) and the EM algorithm-based ProJIVE. Details of the ProJIVE model is currently under review for publication. Here, we use simulated (i.e. toy) data to compare. Our data simulation model is outlined in the ProJIVE manuscript. In order to follow along with the analyses presented herein, please make sure to download the files "Functions_for_PJIVE.R" and "Functions_for_CJIVE.R" from our GitHub page github.com/BenjaminRisk/ProJIVE and save it in the same directory as this document. Both the program to simulate data and ProJIVE are currently designed to handle exactly $K \geq 2$ data sets.

```{r setup, include=FALSE}

require("r.jive")
require("fields")
require("extraDistr")

# NOTE: Change the location of 'prog.dir' to the location where you have saved the file 'Functions_to_SimulateData.R'
prog.dir = "."
source(file.path(prog.dir, "Functions_for_CJIVE.R"))
source(file.path(prog.dir, "Functions_for_PJIVE.R"))
set.seed(0)
```

## Input Parameters to Simulate Data
Input parameters for generating data include the following quantities: $n$ - number of subjects; $p_k$ - number of features in $k^{th}$ data block ($k=1,\dots,K$); $r_J$ - joint rank; $r_{Ik}$ - individual rank for $k^{th}$ data block; $R_{Jk}^2$ - proportion of total variation n $k^{th}$ data block attributable to the joint signal; $R_{Ik}^2$ - proportion of total variation n $k^{th}$ data block attributable to it's individual signal. Note: the last parameters described were proportions and therefore require $R_{Jk}^2 + R_{Ik}^2 \leq 1$ for each $k$. The user may also set a 'seed number' to ensure reproducibility. We use seed number = 0.

```{r, echo = TRUE}
# Sample size
n = 500

# Number of features per block (input as a vector)
p1 = 30
p2 = 50 

# Joint Rank 
r.J = 1

# Individual rank for each block (input as a vector)
r.I1 = 2
r.I2 = 2

# Proportions of joint variation explained (input as a vector)
JntVarEx1 = 0.05 
JntVarEx2 = 0.5

# Proportions of individual variation explained (input as a vector)
IndVarEx1 = 0.25
IndVarEx2 = 0.25
```

### The GenerateToyData Function
In addition to the input parameters defined above, the function "GenerateToyData" has the following parameters: 
'equal.eig' - logical (TRUE/FALSE), which allows the user to specify whether components within a data-blocks joint (or individual) signal should be equally weighted. The default is FALSE

'JntVarAdj' - logical (TRUE/FALSE): Specify whether signal matrices should be weighted to achieve the desired proportions of variation attributable to the joint signal

'SVD.plots' - logical (TRUE/FALSE): Should plots of signal matrix singular values be produced to verify ranks?

'Error' - logical (TRUE/FALSE): Should the data be noise-contaminated?

'print.cor' - logical (TRUE/FALSE), Print the correlation matrix for the scores? (Allows one to assess orthogonality between scores/components)

'Loads' - char: Toy data can be generated with loadings from 'Gaussian', 'Rademacher', or 'Double_Exp'  (double exponential) distributions. Loadings can also be fixed at binary (0/1) values assigned to half of the variables with keyword 'Fixed'

'Scores' - char: Joint subject scores can be randomly generated from 'Gaussian', 'Binomial', or 'Gaussian_Mixture' distributions. The last refers to a  mixture of Gaussians with unit variance, where $20\%$ have mean $-4$, $50\%$ have mean $0$, and $30\%$ have mean $4$. In all cases individual scores are standard Gaussian  

```{r, echo = TRUE}
true_signal_ranks = r.J + c(r.I1,r.I2)                          ##true ranks of overall signals
layout(matrix(1:4, nrow = 2))
ToyDat = GenerateToyData(n = n, p=c(p1, p2), JntVarEx=c(JntVarEx1, JntVarEx2), 
                         IndVarEx = c(IndVarEx1, IndVarEx2), jnt_rank = r.J,
                         equal.eig = F, ind_ranks=c(r.I1, r.I2), JntVarAdj = T, SVD.plots = T,
                         Error = T, print.cor = T, Loads = "Rademacher", Scores = "Gaussian")
layout(1)
```

```{r, echo = TRUE}
blocks <- ToyDat[["Data Blocks"]]

##Setup input parameters to use Gavin's EM code
Y = do.call(cbind, blocks); P = c(p1, p2); Q = c(r.J,r.I1,r.I2)

# Theta
JntScores = ToyDat[['Scores']][['Joint']]
IndivScore.X1 = ToyDat[['Scores']][["Indiv"]][,c("Indiv X1 Score 1", "Indiv X1 Score 2")]
IndivScore.X2 = ToyDat[['Scores']][["Indiv"]][,c("Indiv X2 Score 1", "Indiv X2 Score 2")]

# WJ1 and WJ2
JntLd.X1 = t(ToyDat[['Loadings']][["Joint"]][[1]])
JntLd.X2 = t(ToyDat[['Loadings']][["Joint"]][[2]])

# WI1 and WI2
IndivLd.X1 = t(ToyDat[['Loadings']][["Indiv"]][[1]])
IndivLd.X2 = t(ToyDat[['Loadings']][["Indiv"]][[2]])
```

```{r, echo = FALSE}
# Signal matrices
JX1 = ToyDat[['Data Components']]$JointSignalMatrices[[1]]
JX2 = ToyDat[['Data Components']]$JointSignalMatrices[[2]]
IX1 = ToyDat[['Data Components']]$IndivSignalMatrices[[1]]
IX2 = ToyDat[['Data Components']]$IndivSignalMatrices[[2]]

# Noise Matrices
EX1 = ToyDat[['Data Components']]$NoiseMatrices[[1]]
EX2 = ToyDat[['Data Components']]$NoiseMatrices[[2]]

# Total Signal matrices
AX1 = JX1 + IX1 
AX2 = JX2 + IX2 

## Proportions of variation explained
JVE.X1 = MatVar(JX1)/MatVar(blocks[[1]])
JVE.X2 = MatVar(JX2)/MatVar(blocks[[2]])

IVE.X1 = MatVar(IX1)/MatVar(blocks[[1]])
IVE.X2 = MatVar(IX2)/MatVar(blocks[[2]])

TotVE.X1 = MatVar((JX1 + IX1))/MatVar(blocks[[1]])
TotVE.X2 = MatVar((JX2 + IX2))/MatVar(blocks[[2]])

```

Check that values of $R_{Jk}^2$ and $R_{Ik}^2$ in the simulated data match the desired quantities.
```{r, echo = FALSE}
bquote(paste(R[J1]^2, '=' , .(round(JVE.X1,4))))
bquote(paste(R[J2]^2, '=' , .(round(JVE.X2,4))))
bquote(paste(R[I1]^2, '=' , .(round(IVE.X1,4))))
bquote(paste(R[I2]^2, '=' , .(round(IVE.X2,4))))
```

# JIVE analyses
## ProJIVE
Now, compute the ProJIVE solution and graph joint subject scores (not individual scores) in addition to joint and individual loadings. The ProJIVE algorithm requires an initial guess for the noise variances and the variable loadings. By default, these values are chosen at random from Gaussian distributions. The initial values can also be chosen by the user. Initial noise error variances can be entered as a vector of length $K$ and initial loadings as a list of length $K$ with the $k^{th}$ entry as a $p_k \times (r_J + r_{Ik})$ matrix of real numbers. Alternatively, we recommend the settings below. A user can initialize the noise variances at their MLE from the probabilistic model of principal components analysis (attributed to Tipping and Bishop) by setting 'sig_hat = "MLE"' and variable loadings at their AJIVE (i.e. CJIVE for $K=2$) solution with 'init.loads = "AJIVE"'. Lastly, ProJIVE automatically generates graphs of congvergence criterion for each iteration of the EM algorithm. These can be turned off by specifying 'plots = FALSE' in the ProJIVE call.   
```{r, echo = TRUE}
##### ProJIVE
PJIVE.res = ProJIVE_EM(Y=Y, P=P, Q=Q, Max.iter=10000, diff.tol=1e-7, sig_hat = "MLE", init.loads = "AJIVE")

PJIVE.scores = PJIVE.res$SubjectScoreMatrix
PJIVE.loads.X1 = PJIVE.res$LoadingMatrix[1:p1,-(sum(Q):(sum(Q[-3])+1))]
PJIVE.loads.X2 = PJIVE.res$LoadingMatrix[-(1:p1),-(r.J+1:r.I1)]
```

```{r, echo = FALSE}
layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
plot(JntScores, PJIVE.scores[,1:r.J], xlab = "True Joint Scores", ylab = "ProJIVE Joint Scores", 
     col = c(rep("orange",n), rep("green",n), rep("purple",n)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, PJIVE.scores[,1:r.J]), 3)))
plot(IndivScore.X1, PJIVE.scores[,r.J+(1:r.I1)], 
     xlab = "True Indiv X1 Scores", ylab = "ProJIVE Indiv X1 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X1, PJIVE.scores[,r.J+(1:r.I1)]), 3)))
plot(IndivScore.X2, PJIVE.scores[,r.J+r.I1+(1:r.I2)], 
     xlab = "True Indiv X2 Scores", ylab = "ProJIVE Indiv X2 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X2, PJIVE.scores[,r.J+r.I1+(1:r.I2)]), 3)))
plot.new(); title("ProJIVE Results")
legend("left", paste("Comp.", 1:2), pch = 1, col  = c("orange", "green"),bty = "n" )
plot(JntLd.X1, PJIVE.loads.X1[,1:r.J], xlab = "True Joint Loadings X", ylab = "ProJIVE Joint Loadings X", 
     col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X1, PJIVE.loads.X1[,1:r.J]), 3)))
plot(JntLd.X2, PJIVE.loads.X2[,1:r.J], xlab = "True Joint Loadings X2", ylab = "ProJIVE Joint Loadings X2", 
     col = c(rep("orange",p2), rep("green",p2), rep("purple",p2)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X2, PJIVE.loads.X2[,1:r.J]), 3)))
plot(IndivLd.X1, PJIVE.loads.X1[,-(1:r.J)], xlab = "True Individual Loadings X", ylab = "ProJIVE Individual Loadings X", 
     col = c(rep("orange",p1), rep("green",p1)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X1, PJIVE.loads.X1[,-(1:r.J)]), 3)))
plot(IndivLd.X2, PJIVE.loads.X2[,-(1:r.J)], xlab = "True Individual Loadings X2", ylab = "ProJIVE Individual Loadings X2", 
     col = c(rep("orange",p2), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X2, PJIVE.loads.X2[,-(1:r.J)]), 3)))
layout(1)
```

## CJIVE
Now, compute the CJIVE solution and graph subject scores and variable loadings.
```{r, echo = TRUE}
#### CJIVE
CJIVE.res = cc.jive(blocks, true_signal_ranks, r.J, perm.test = FALSE)

# CJIVE signal matrix estimates
J.hat = CJIVE.res$sJIVE$joint_matrices
I.hat = CJIVE.res$sJIVE$indiv_matrices

# CJIVE loading estimates
CJIVE.IndScores = lapply(I.hat, function(x) x[['u']])

# CJIVE loading estimates
WJ = lapply(J.hat, function(x) x[['v']])
WI = lapply(I.hat, function(x) x[['v']])
```

```{r, echo = FALSE}
# Plots of CJIVE estimates against true counterparts
layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
plot(JntScores, CJIVE.res$CanCorRes$Jnt_Scores, xlab = "True Joint Scores",ylab = "CJIVE Joint Scores", 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, CJIVE.res$CanCorRes$Jnt_Scores), 3)))
plot(IndivScore.X1, CJIVE.IndScores[[1]], 
     xlab = "True Indiv X1 Scores", ylab = "CJIVE Indiv X1 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X1, CJIVE.IndScores[[1]]), 3)))
plot(IndivScore.X2, CJIVE.IndScores[[2]], 
     xlab = "True Indiv X2 Scores", ylab = "CJIVE Indiv X2 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X2, CJIVE.IndScores[[2]]), 3)))
plot.new(); title("CJIVE Results")
legend("left", paste("Comp.", 1:2), pch = 1, col  = c("orange", "green"),bty = "n" )
plot(JntLd.X1, WJ[[1]], xlab = "True Joint Loadings X", ylab = "CJIVE Joint Loadings X", 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X1, WJ[[1]]), 3)))
plot(JntLd.X2, WJ[[2]], xlab = "True Joint Loadings X2", ylab = "CJIVE Joint Loadings X2", 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X2, WJ[[2]]), 3)))
plot(IndivLd.X1, WI[[1]], xlab = "True Individual Loadings X", ylab = "CJIVE Individual Loadings X", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X1, WI[[1]]), 3)))
plot(IndivLd.X2, WI[[2]], xlab = "True Individual Loadings X2", ylab = "CJIVE Individual Loadings X2", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X2, WI[[2]]), 3)))
layout(1)
```

# Gaussian Mixture Scores: Toy data which do not correspond to the ProJIVE model
When study participants can be categorized due to an exogenous variable that is associated with the joint sources of variation (i.e. diagnosis categories, etc.), joint subject scores may be used to categorize participants. The code snippet here generates toy data that reflect this scenario by taking joint subject scores from Gaussian mixture model.

```{r, echo = TRUE}
true_signal_ranks = r.J + c(r.I1,r.I2)                          ##true ranks of overall signals
layout(matrix(1:4, nrow = 2))
ToyDat = GenerateToyData(n = n, p = c(p1, p2), JntVarEx = c(JntVarEx1, JntVarEx2), 
                         IndVarEx = c(IndVarEx1, IndVarEx2), jnt_rank = r.J,
                         equal.eig = F, ind_rank = c(r.I1, r.I2), JntVarAdj = T, SVD.plots = T,
                         Error = T, print.cor = F, Loads = "Double_Exp", Scores = "Gaussian_Mixture")
layout(1)
```

```{r, echo = TRUE}
blocks <- ToyDat[[2]]

##Setup input parameters to use Gavin's EM code
Y = do.call(cbind, blocks); P = c(p1, p2); Q = c(r.J,r.I1,r.I2)

# Theta
JntScores = ToyDat[['Scores']][['Joint']]
IndivScore.X1 = ToyDat[['Scores']][["Indiv"]][,c("Indiv X1 Score 1", "Indiv X1 Score 2")]
IndivScore.X2 = ToyDat[['Scores']][["Indiv"]][,c("Indiv X2 Score 1", "Indiv X2 Score 2")]

# WJ1 and WJ2
JntLd.X1 = t(ToyDat[['Loadings']][["Joint"]][[1]])
JntLd.X2 = t(ToyDat[['Loadings']][["Joint"]][[2]])

# WI1 and WI2
IndivLd.X1 = t(ToyDat[['Loadings']][["Indiv"]][[1]])
IndivLd.X2 = t(ToyDat[['Loadings']][["Indiv"]][[2]])
```

```{r, echo = FALSE}
JX1 = ToyDat[['Data Components']]$JointSignalMatrices[[1]]
JX2 = ToyDat[['Data Components']]$JointSignalMatrices[[2]]
IX1 = ToyDat[['Data Components']]$IndivSignalMatrices[[1]]
IX2 = ToyDat[['Data Components']]$IndivSignalMatrices[[2]]

# Noise Matrices
EX1 = ToyDat[['Data Components']]$NoiseMatrices[[1]]
EX2 = ToyDat[['Data Components']]$NoiseMatrices[[2]]

## Proportions of variation explained
JVE.X1 = MatVar(JX1)/MatVar(blocks[[1]])
JVE.X2 = MatVar(JX2)/MatVar(blocks[[2]])

IVE.X1 = MatVar(IX1)/MatVar(blocks[[1]])
IVE.X2 = MatVar(IX2)/MatVar(blocks[[2]])

TotVE.X1 = MatVar((JX1 + IX1))/MatVar(blocks[[1]])
TotVE.X2 = MatVar((JX2 + IX2))/MatVar(blocks[[2]])

```

Check that values of $R_{Jk}^2$ and $R_{Ik}^2$ in the simulated data match the desired quantities.
```{r, echo = FALSE}
bquote(paste(R[J1]^2, '=' , .(round(JVE.X1,4))))
bquote(paste(R[J2]^2, '=' , .(round(JVE.X2,4))))
bquote(paste(R[I1]^2, '=' , .(round(IVE.X1,4))))
bquote(paste(R[I2]^2, '=' , .(round(IVE.X2,4))))
```

## CJIVE
Now, compute the CJIVE solution and graph subject scores and loadings.
```{r, echo = TRUE}
CJIVE.res = cc.jive(blocks, true_signal_ranks, r.J, perm.test = FALSE)

# CJIVE signal matrix estimates
J.hat = CJIVE.res$sJIVE$joint_matrices
I.hat = CJIVE.res$sJIVE$indiv_matrices

# CJIVE loading estimates
CJIVE.IndScores = lapply(I.hat, function(x) x[['u']])

# CJIVE loading estimates
WJ = lapply(J.hat, function(x) x[['v']])
WI = lapply(I.hat, function(x) x[['v']])
```

```{r, echo = FALSE}
# Plots of CJIVE estimates against true counterparts
layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
plot(JntScores, CJIVE.res$CanCorRes$Jnt_Scores, xlab = "True Joint Scores",ylab = "CJIVE Joint Scores", 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, CJIVE.res$CanCorRes$Jnt_Scores), 3)))
plot(IndivScore.X1, CJIVE.IndScores[[1]], 
     xlab = "True Indiv X1 Scores", ylab = "CJIVE Indiv X1 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X1, CJIVE.IndScores[[1]]), 3)))
plot(IndivScore.X2, CJIVE.IndScores[[2]], 
     xlab = "True Indiv X2 Scores", ylab = "CJIVE Indiv X2 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X2, CJIVE.IndScores[[2]]), 3)))
plot.new(); title("CJIVE Results")
legend("left", paste("Comp.", 1:2), pch = 1, col  = c("orange", "green"),bty = "n" )
plot(JntLd.X1, WJ[[1]], xlab = "True Joint Loadings X", ylab = "CJIVE Joint Loadings X", 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X1, WJ[[1]]), 3)))
plot(JntLd.X2, WJ[[2]], xlab = "True Joint Loadings X2", ylab = "CJIVE Joint Loadings X2", 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X2, WJ[[2]]), 3)))
plot(IndivLd.X1, WI[[1]], xlab = "True Individual Loadings X", ylab = "CJIVE Individual Loadings X", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X1, WI[[1]]), 3)))
plot(IndivLd.X2, WI[[2]], xlab = "True Individual Loadings X2", ylab = "CJIVE Individual Loadings X2", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X2, WI[[2]]), 3)))
layout(1)
```

## ProJIVE
Now, compute the PJIVE solution and graph subject scores and loadings.
```{r, echo = TRUE}
##### ProJIVE
PJIVE.res = ProJIVE_EM(Y=Y, P=P, Q=Q, Max.iter=10000, diff.tol=1e-7, sig_hat = "MLE", init.loads = "AJIVE")

PJIVE.scores = PJIVE.res$SubjectScoreMatrix
PJIVE.loads.X1 = PJIVE.res$LoadingMatrix[1:p1,-(sum(Q):(sum(Q[-3])+1))]
PJIVE.loads.X2 = PJIVE.res$LoadingMatrix[-(1:p1),-(r.J+1:r.I1)]
```

```{r, echo = FALSE}
layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
plot(JntScores, PJIVE.scores[,1:r.J], xlab = "True Joint Scores", ylab = "ProJIVE Joint Scores", 
     col = c(rep("orange",n), rep("green",n), rep("purple",n)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, PJIVE.scores[,1:r.J]), 3)))
plot(IndivScore.X1, PJIVE.scores[,r.J+(1:r.I1)], 
     xlab = "True Indiv X1 Scores", ylab = "ProJIVE Indiv X1 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X1, PJIVE.scores[,r.J+(1:r.I1)]), 3)))
plot(IndivScore.X2, PJIVE.scores[,r.J+r.I1+(1:r.I2)], 
     xlab = "True Indiv X2 Scores", ylab = "ProJIVE Indiv X2 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X2, PJIVE.scores[,r.J+r.I1+(1:r.I2)]), 3)))
plot.new(); title("ProJIVE Results")
legend("left", paste("Comp.", 1:2), pch = 1, col  = c("orange", "green"),bty = "n" )
plot(JntLd.X1, PJIVE.loads.X1[,1:r.J], xlab = "True Joint Loadings X", ylab = "ProJIVE Joint Loadings X", 
     col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X1, PJIVE.loads.X1[,1:r.J]), 3)))
plot(JntLd.X2, PJIVE.loads.X2[,1:r.J], xlab = "True Joint Loadings X2", ylab = "ProJIVE Joint Loadings X2", 
     col = c(rep("orange",p2), rep("green",p2), rep("purple",p2)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X2, PJIVE.loads.X2[,1:r.J]), 3)))
plot(IndivLd.X1, PJIVE.loads.X1[,-(1:r.J)], xlab = "True Individual Loadings X", ylab = "ProJIVE Individual Loadings X", 
     col = c(rep("orange",p1), rep("green",p1)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X1, PJIVE.loads.X1[,-(1:r.J)]), 3)))
plot(IndivLd.X2, PJIVE.loads.X2[,-(1:r.J)], xlab = "True Individual Loadings X2", ylab = "ProJIVE Individual Loadings X2", 
     col = c(rep("orange",p2), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X2, PJIVE.loads.X2[,-(1:r.J)]), 3)))
layout(1)
```

We can also check the histogram of joint scores from each method for the multimodality that arises from Gaussian mixtures and check that there is a statistically significant association between joint scores and group assignment based on the miture model. 
```{r, echo = TRUE}
## Proportions of groups for mixture
mix.probs = c(0.2, 0.5, 0.3)
n.mix = n*mix.probs
mix.group = factor(c(rep(1, n.mix[1]),rep(2, n.mix[2]),rep(3, n.mix[3])))
JntScores.dat = data.frame(V1 = c(CJIVE.res$CanCorRes$Jnt_Scores, PJIVE.scores[,1:r.J], JntScores))
JntScores.dat$group = factor(c(mix.group, mix.group, mix.group), labels = paste("Group", 1:3))
JntScores.dat$Method = factor(rep(1:3, each = n), labels = c("CJIVE", "ProJIVE", "True Scores"))
ggplot(data = JntScores.dat, aes(V1, fill = group)) + geom_histogram(bins = 20, position = "dodge") + 
  facet_wrap(Method~., scales = "free") + labs(x = element_blank())
```
