---
title: "Problematic Identifiability Example"
author: "Benjamin Risk"
output: html_document
---

## Linear dependence between columns of $W_{Jk}$ and $W_{Ik}$ and identifiability 

Here, we present an example that shows issues that arise when 
$W_{Jk}$ and $W_{Ik}$ are not linearly independent:

   i) the joint versus individual variation is not identifiable, 
   
   ii) the individual components are not identifiable,
   
   iii) the number of individual components is not identifiable. 

To examine the invariance of the likelihood, it suffices to consider the covariance matrices. Recall $Cov(x_{ik}) = W_{Jk}W_{Jk}^\top + W_{Ik}W_{Ik}^\top + \sigma_k^2 I$. We can ignored $\sigma_k^2$ in the account that follows. 

Generate loadings:
```{r}
WJ1 = diag(5)[,c(1,2)]
WI1 = diag(5)[,c(3,4)]
```


For dataset 2, we will create a non-zero subspace in the  $col(W_{J2}) \cap col(W_{I2})$:

```{r}
WJ2 = diag(5)[,c(1,2)]
WI2 = diag(5)[,c(1,3)]
```

Consider the setting described in theorem 1. Let $T_2 = (T_1^{-1})^\top$. In order for the individual components of the alternative decomposition to be well-defined, their covariance matrix $W_{Ik}W_{Ik}^\top + W_{Jk}(I - T_k T_k^\top)W_{Jk}^\top$ needs to be positive semidefinite. If $col(W_{J2}) \cap col(W_{I2}) \ne \{0\}$, an eigenvalue of $W_{J2}(I-T_2 T_2^\top)W_{J2}^\top$ can be negative if it is offset by the energy in the corresponding direction in $W_{I2}$. 

```{r}
T1 = diag(c(sqrt(0.5+0.1),1)) 
T2 = t(solve(T1))
```

Construct the new joint loadings matrices:

```{r}
WJ1star = WJ1%*%T1
WJ2star = WJ2%*%T2
```

Construct the new individual components covariance matrix:

```{r}
covWI1star = WI1%*%t(WI1) + WJ1%*%(diag(ncol(WJ1))-T1%*%t(T1))%*%t(WJ1)
eig.WI1star = eigen(covWI1star)
#note no negative eigenvalues:
eig.WI1star$values
(rI1.star = sum(eig.WI1star$values>1e-8))
WI1star = eig.WI1star$vectors[,1:rI1.star]%*%diag(sqrt(eig.WI1star$values)[1:rI1.star])
```

We now have THREE non-zero eigenvalues, i.e., the number of individual components is not identifiable. This occurs because the overlap in the column spaces of $W_{J2}$ and $W_{I2}$ allows $(I - T_2 T_2^\top)$ to have negative eigenvalues ($T_2 = (T_1^{-1})^\top$), and since $I - T_1 T_1^\top$ is positive definite, $W_{J1}(I - T_1 T_1^\top) W_{J1}^\top$ has a non-negligible contribution of $W_{J1}$. 

Next, define the new loadings of the individual components of dataset 2:
```{r}
covWI2star = WI2%*%t(WI2) + WJ2%*%(diag(ncol(WJ2))-T2%*%t(T2))%*%t(WJ2)
eig.WI2star = eigen(covWI2star)
# check all positive:
eig.WI2star$values
# we have a psd matrix, indicating a valid choice of T1

WI2star = eig.WI2star$vectors[,1:2]%*%diag(sqrt(eig.WI2star$values)[1:2])
```

Finally, check covariance matrices are equivalent, which implies the likelihoods are equivalent:
```{r}
cov1 = WJ1%*%t(WJ1)+WI1%*%t(WI1)
cov2 = WJ2%*%t(WJ2)+WI2%*%t(WI2)

cov1star = WJ1star%*%t(WJ1star)+WI1star%*%t(WI1star)
cov2star = WJ2star%*%t(WJ2star)+WI2star%*%t(WI2star)
norm(cov1-cov1star,type = 'F')
norm(cov2-cov2star,type = 'F')

covX1X2 = WJ1%*%t(WJ2)
covX1X2star = WJ1star%*%t(WJ2star)
norm(covX1X2-covX1X2star,type='F')
```

This demonstrates a ProJIVE model in which the joint versus individual variation is not identifiable, nor are the individual components identifiable. 

## Identifiability up to orthogonal rotations

Hypothesis: $W_{Jk}$ and $W_{Ik}$ are identifiable up to orthogonal transformations when the columns of $W_{Jk}$ and $W_{Ik}$ are linearly independent

Here is an example to see how the negative eigenvalues can arise, which implies that the $T_1$ matrix does not satisfy the constraints $W_{Ik}W_{Ik}^\top + W_{Jk}(I - T_k T_k^\top)W_{Jk}^\top \succeq 0$ for  $T_2 = (T_1^{-1})^\top$.

```{r}
WJ1 = matrix(c(1,1,1,0,0),ncol=1)
WI1 = matrix(c(1,2,0,0,0),ncol=1)

# columns are linearly independent:
svd(cbind(WJ1,WI1))$d


T1 = 1.01

# individual components covariance matrix:
covWI1star = WI1%*%t(WI1) + WJ1%*%(diag(ncol(WJ1))-T1%*%t(T1))%*%t(WJ1)
# it is non positive semidefinite:
eigen(covWI1star)$values
```


Here is another example where linear independence of the columns of $W_{J1}$ and $W_{I1}$ appears to be sufficient to restricting the class of allowable T1 to orthogonal matrices. We tried to find a counterexample to the claim that linear independence of columns of $W_{Jk}$ and $W_{Ik}$ is necessary and sufficient. Specifically, the negative eigenspace of $\bW_{J1}^\top (I - T_1 T_1^\top) W_{J1}$ is entirely contained in the projection of $W_{I1}$ to the negative eigenspace. Nonetheless, we are still unable to "swamp" the negative eigenvalue by $W_{I1} W_{I1}^\top$. In other words, $T_1$ appears to be constrained to the class of orthogonal matrices, and the joint and individual components are identifiable up to signed rotation. 
```{r}
WI1 = matrix(c(1,0.5,0.5),nrow=3)
WJ1 = matrix(c(1,0,0,0,1,0),nrow=3)
svd(cbind(WJ1,WI1))$d

T1=diag(c(1,sqrt(1.5)))
L = WJ1%*%(diag(2)-T1%*%t(T1))%*%t(WJ1)
Psi = 100*WI1%*%t(WI1)+L
eigPsi = eigen(Psi)
eigPsi$values
# eigPsi is not PSD

# projection of WI1 onto negative eigenspace of WJ1
# spans negative eigenspace of WJ1. Why can't I make the eigenvalue negative
u1=matrix(c(0,1,0),nrow=3)
t(u1)%*%Psi%*%u1

# interestingly, the negative eigenvector can have a positive value in the quadratic form. 
```

