---
title: "Examining Mixed Joint Variation in Simulated Data for ProJIVE"
author: "Raphiel J. Murden"
date: "3/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
prog.dir = "H:/My Documents/P-JIVE/Programs/Functions"
prog.gipca.dir = "H:/My Documents/P-JIVE/Programs/GeneralizedIntegrativePCA-master/Functions"
source(file.path(prog.dir, "Functions_for_CJIVE.R"))
source(file.path(prog.dir, "Functions_for_PJIVE.R"))
# source(file.path(prog.dir, "Functions_for_PJIVE_DEBUG.R"))
gipca.files = list.files(prog.gipca.dir,full.names = TRUE)
for(file.nm in gipca.files){source(file.nm)}
doc.dir = "H:/My Documents/P-JIVE/Programs/Examples/Output"

rep_number = 217
r.J = 3
r.I1 = 2
r.I2 = 2
#outdir = args[2]
n = 1000
p1 = 20
p2 = 20 ####Note that p1 and p2 differ when compared to values used in simulations
JntVarEx1 = 0.5
JntVarEx2 = 0.05
#files = list.files(outdir)
IndVarEx1 = 0.25
IndVarEx2 = 0.25

set.seed(rep_number) ##To ensure that any randomized processes give the same results each time
```

## Introduction

Examining the issue found in simulation studies of analyses when joint variance is mixed across data sets but the number of features is the same across data sets. So far, findings in this single trial reflect the trends in simulation studies; namely when first joint variation is small, results are less accurate. This violates the idea that the method is symmetric and the order in which data sets are entered should not make a difference. This finding appears for both R.JIVE and ProJIVE.

To examine this issue data are generated using parameters from the ProJIVE manuscript so that joint scores are drawn from a Gaussian Mixture distribution, individual scores are drawn from standard Gaussian distributions, loadings are drawn from Rademacher distributions, the sample size is $n=1000$, the number of features are $p_1=p_2=20$, the individual proportions of variance explained $R_{I1}^2=R_{I2}^2=0.25$, individual ranks $r_{I1}=r_{I2}=2$, and joint rank is $r_J=3$. The proportions of joint variance explained in each data set are $R_{J1}^2=0.5$ and $R_{J2}^2=0.05$ for the first example. Then the data sets labels are switched so that $R_{J1}^2=0.05$ and $R_{J2}^2=0.5$. By using the same data sets and simply changing their labels, we show that neither R.JIVE nor ProJIVE is symmetric with respect to data set labels. Specifically, this vignette seems to suggest that both methods are less accurate when the joint proportion of variance in the first data set is much smaller than that in the second data set. 

The first two set of plots shows results of analyses when data are generated so that $R_{J1}^2=0.5$ and $R_{J2}^2=0.05$.

```{r, echo=FALSE}
true_signal_ranks = r.J + c(r.I1,r.I2)                          ##true ranks of overall signals
ToyDat = GenerateToyData(n = n, p1 = p1, p2 = p2, JntVarEx1 = JntVarEx1, JntVarEx2 = JntVarEx2, 
                         IndVarEx1 = IndVarEx1, IndVarEx2 =  IndVarEx2, jnt_rank = r.J,
                         equal.eig = T,ind_rank1 = r.I1, ind_rank2 = r.I2, JntVarAdj = T, SVD.plots = F,
                         Error = T, print.cor = F, Loads = "Rademacher", Scores = "Gaussian_Mixture")
blocks <- ToyDat[[2]]
##Setup input parameters to use Gavin's EM code
Y = do.call(cbind, blocks); P = c(p1, p2); Q = c(r.J,r.I1,r.I2)

# Theta
JntScores = ToyDat[['Scores']][['Joint']]
IndivScore.X = ToyDat[['Scores']][["Indiv_1"]]
IndivScore.Y = ToyDat[['Scores']][["Indiv_2"]]

# WJ1 and WJ2
JntLd.X = t(ToyDat[['Loadings']][["Joint_1"]])
JntLd.Y = t(ToyDat[['Loadings']][["Joint_2"]])

# WI1 and WI2
IndivLd.X =t(ToyDat[['Loadings']][["Indiv_1"]])
IndivLd.Y = t(ToyDat[['Loadings']][["Indiv_2"]])

# Signal matrices
JX = ToyDat[[1]]$J1
JY = ToyDat[[1]]$J2
IX = ToyDat[[1]]$I1
IY = ToyDat[[1]]$I2

# Noise Matrices
EX = ToyDat[[1]]$E1
EY = ToyDat[[1]]$E2

# Total Signal matrices
AX = JX + IX
AY = JY + IY

## Proportions of variation explained
JVE.X = MatVar(JX)/MatVar(blocks[[1]])
JVE.Y = MatVar(JY)/MatVar(blocks[[2]])

IVE.X = MatVar(IX)/MatVar(blocks[[1]])
IVE.Y = MatVar(IY)/MatVar(blocks[[2]])

TotVE.X = MatVar((JX + IX))/MatVar(blocks[[1]])
TotVE.Y = MatVar((JY + IY))/MatVar(blocks[[2]])

#############################################
############### ProJIVE  ####################
# pdf(file = file.path(doc.dir, "InitialValueInvestigation.pdf"))
PJIVE.res = ProJIVE_EM(Y=Y, P=P, Q=Q, diff.tol=1e-5, plots = FALSE, sig_hat = "MLE", init.loads = "CJIVE")

PJIVE.scores = PJIVE.res$SubjectScoreMatrix
PJIVE.loads.X = PJIVE.res$LoadingMatrix[1:p1,-(sum(Q):(sum(Q[-3])+1))]
PJIVE.loads.Y = PJIVE.res$LoadingMatrix[-(1:p1),-(r.J+1:r.I1)]

layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
plot(JntScores, PJIVE.scores[,1:r.J], xlab = "True Joint Scores", ylab = "ProJIVE Joint Scores", 
     col = c(rep("orange",n), rep("green",n), rep("purple",n)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, PJIVE.scores[,1:r.J]), 3)))
plot(IndivScore.X, PJIVE.scores[,4:5], xlab = "True Indiv 1 Scores", ylab = "ProJIVE Indiv 1 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X, PJIVE.scores[,4:5]), 3)))
plot(IndivScore.Y, PJIVE.scores[,6:7], xlab = "True Indiv 2 Scores", ylab = "ProJIVE Indiv 2 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.Y, PJIVE.scores[,6:7]), 3)))
plot.new(); title("ProJIVE Results with \n R_J1^2=0.5")
legend("left", paste("Comp.", 1:3), pch = 1, col  = c("orange", "green", "purple"),bty = "n" )
plot(JntLd.X, PJIVE.loads.X[,1:r.J], xlab = "True Joint Loadings X", ylab = "ProJIVE Joint Loadings X", 
     col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X, PJIVE.loads.X[,1:r.J]), 3)))
plot(JntLd.Y, PJIVE.loads.Y[,1:r.J], xlab = "True Joint Loadings Y", ylab = "ProJIVE Joint Loadings Y", 
     col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.Y, PJIVE.loads.Y[,1:r.J]), 3)))
plot(IndivLd.X, PJIVE.loads.X[,-(1:r.J)], xlab = "True Individual Loadings X", ylab = "ProJIVE Individual Loadings X", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X, PJIVE.loads.X[,-(1:r.J)]), 3)))
plot(IndivLd.Y, PJIVE.loads.Y[,-(1:r.J)], xlab = "True Individual Loadings Y", ylab = "ProJIVE Individual Loadings Y", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.Y, PJIVE.loads.Y[,-(1:r.J)]), 3)))
layout(1)

# blocks.t = lapply(blocks, t)
# R.JIVE.res = jive(blocks.t, rankJ = r.J, rankA = true_signal_ranks - r.J, method = "given")
# 
# # R.JIVE signal matrix estimates
# J.hat = lapply(R.JIVE.res$joint, svd, nu = r.J, nv = r.J)
# I.hat = list()
# I.hat[[1]] = svd(R.JIVE.res$individual[[1]], nu = r.I1, nv = r.I1)
# I.hat[[2]] = svd(R.JIVE.res$individual[[2]], nu = r.I2, nv = r.I2)
# 
# # R.JIVE Subject Scores
# R.JIVE.jnt.scores = svd(do.call(rbind, R.JIVE.res$joint), nu = r.J, nv = r.J)$v
# R.JIVE.ind.scores.X = I.hat[[1]][['v']]
# R.JIVE.ind.scores.Y = I.hat[[2]][['v']]
# 
# # R.JIVE loading estimates
# WJ = lapply(J.hat, function(x) x[['u']])
# WI = lapply(I.hat, function(x) x[['u']])
# 
# # Plots of R.JIVE estimates against true counterparts
# # layout(matrix(c(1,1,1:7), 3, byrow = TRUE), heights = c(.1, 0.4, 0.4))
# # plot(1, type = "n", axes=FALSE, xlab="", ylab=""); title(sub = "temp")
# layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
# plot(JntScores, R.JIVE.jnt.scores, xlab = "True Joint Scores",ylab = "R.JIVE Joint Scores", 
#      col = c(rep("orange",n), rep("green",n), rep("purple",n)),  
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, R.JIVE.jnt.scores), 3)))
# plot(IndivScore.X, R.JIVE.ind.scores.X, xlab = "True Indiv 1 Scores", ylab = "ProJIVE Indiv 1 Scores", 
#      col = c(rep("orange",n), rep("green",n)),
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X, R.JIVE.ind.scores.X), 3)))
# plot(IndivScore.Y, R.JIVE.ind.scores.Y, xlab = "True Indiv 2 Scores", ylab = "ProJIVE Indiv 2 Scores", 
#      col = c(rep("orange",n), rep("green",n)),
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.Y, R.JIVE.ind.scores.Y), 3)))
# # plot(1,lwd=0,axes=F,xlab="",ylab="", "n")
# plot.new(); title("R.JIVE Results with \n R_J1^2=0.5")
# legend("left", paste("Comp.", 1:3), pch = 1, col  = c("orange", "green", "purple"),bty = "n" )
# plot(JntLd.X, WJ[[1]], xlab = "True Joint Loadings X", ylab = "R.JIVE Joint Loadings X", 
#      col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X, WJ[[1]]), 3)))
# plot(JntLd.Y, WJ[[2]], xlab = "True Joint Loadings Y", ylab = "R.JIVE Joint Loadings Y", 
#      col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.Y, WJ[[2]]), 3)))
# plot(IndivLd.X, WI[[1]], xlab = "True Individual Loadings X", ylab = "R.JIVE Individual Loadings X", 
#      col = c(rep("orange",p1), rep("green",p2)), 
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X, WI[[1]]), 3)))
# plot(IndivLd.Y, WI[[2]], xlab = "True Individual Loadings Y", ylab = "R.JIVE Individual Loadings Y", 
#      col = c(rep("orange",p1), rep("green",p2)), 
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.Y, WI[[2]]), 3)))
# layout(1)
```

## Reverse proportions of variance exaplined by switching labels on data sets

Next we see results of analyses when data labels have been switched so that $R_{J1}^2=0.05$ and $R_{J2}^2=0.5$.

```{r pressure, echo=FALSE}
temp=blocks
blocks=list(temp[[2]],temp[[1]])
##Setup input parameters to use Gavin's EM code
Y = do.call(cbind, blocks); P = c(p1, p2); Q = c(r.J,r.I1,r.I2)

# Theta
JntScores = ToyDat[['Scores']][['Joint']]
IndivScore.X = ToyDat[['Scores']][["Indiv_2"]]
IndivScore.Y = ToyDat[['Scores']][["Indiv_1"]]

# WJ1 and WJ2
JntLd.X = t(ToyDat[['Loadings']][["Joint_2"]])
JntLd.Y = t(ToyDat[['Loadings']][["Joint_1"]])

# WI1 and WI2
IndivLd.X =t(ToyDat[['Loadings']][["Indiv_2"]])
IndivLd.Y = t(ToyDat[['Loadings']][["Indiv_1"]])

# Signal matrices
JX = ToyDat[[1]]$J2
JY = ToyDat[[1]]$J1
IX = ToyDat[[1]]$I2
IY = ToyDat[[1]]$I1

# Noise Matrices
EX = ToyDat[[1]]$E2
EY = ToyDat[[1]]$E1

# Total Signal matrices
AX = JX + IX
AY = JY + IY

## Proportions of variation explained
JVE.X = MatVar(JX)/MatVar(blocks[[2]])
JVE.Y = MatVar(JY)/MatVar(blocks[[1]])

IVE.X = MatVar(IX)/MatVar(blocks[[2]])
IVE.Y = MatVar(IY)/MatVar(blocks[[1]])

TotVE.X = MatVar((JX + IX))/MatVar(blocks[[2]])
TotVE.Y = MatVar((JY + IY))/MatVar(blocks[[1]])

##Looks like the variances are about equal to me ( +/- 10%)?
# plot(apply((blocks[[1]] - AX), 2, sd))
# plot(apply((blocks[[2]] - AY), 2, sd))

#############################################
############### ProJIVE  ####################
# pdf(file = file.path(doc.dir, "InitialValueInvestigation.pdf"))
PJIVE.res = ProJIVE_EM(Y=Y, P=P, Q=Q, diff.tol=1e-5, plots = FALSE, sig_hat = "MLE", init.loads = "CJIVE")

PJIVE.scores = PJIVE.res$SubjectScoreMatrix
PJIVE.loads.X = PJIVE.res$LoadingMatrix[1:p1,-(sum(Q):(sum(Q[-3])+1))]
PJIVE.loads.Y = PJIVE.res$LoadingMatrix[-(1:p1),-(r.J+1:r.I1)]

layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
plot(JntScores, PJIVE.scores[,1:r.J], xlab = "True Joint Scores", ylab = "ProJIVE Joint Scores", 
     col = c(rep("orange",n), rep("green",n), rep("purple",n)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, PJIVE.scores[,1:r.J]), 3)))
plot(IndivScore.X, PJIVE.scores[,4:5], xlab = "True Indiv 1 Scores", ylab = "ProJIVE Indiv 1 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X, PJIVE.scores[,4:5]), 3)))
plot(IndivScore.Y, PJIVE.scores[,6:7], xlab = "True Indiv 2 Scores", ylab = "ProJIVE Indiv 2 Scores", 
     col = c(rep("orange",n), rep("green",n)),
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.Y, PJIVE.scores[,6:7]), 3)))
plot.new(); title("ProJIVE Results with \n R_J1^2=0.05")
legend("left", paste("Comp.", 1:3), pch = 1, col  = c("orange", "green", "purple"),bty = "n" )
plot(JntLd.X, PJIVE.loads.X[,1:r.J], xlab = "True Joint Loadings X", ylab = "ProJIVE Joint Loadings X", 
     col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X, PJIVE.loads.X[,1:r.J]), 3)))
plot(JntLd.Y, PJIVE.loads.Y[,1:r.J], xlab = "True Joint Loadings Y", ylab = "ProJIVE Joint Loadings Y", 
     col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.Y, PJIVE.loads.Y[,1:r.J]), 3)))
plot(IndivLd.X, PJIVE.loads.X[,-(1:r.J)], xlab = "True Individual Loadings X", ylab = "ProJIVE Individual Loadings X", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X, PJIVE.loads.X[,-(1:r.J)]), 3)))
plot(IndivLd.Y, PJIVE.loads.Y[,-(1:r.J)], xlab = "True Individual Loadings Y", ylab = "ProJIVE Individual Loadings Y", 
     col = c(rep("orange",p1), rep("green",p2)), 
     sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.Y, PJIVE.loads.Y[,-(1:r.J)]), 3)))
layout(1)


########################################
##############  R.JIVE   ###############
# blocks.t = lapply(blocks, t)
# R.JIVE.res = jive(blocks.t, rankJ = r.J, rankA = true_signal_ranks - r.J, method = "given")
# 
# # R.JIVE signal matrix estimates
# J.hat = lapply(R.JIVE.res$joint, svd, nu = r.J, nv = r.J)
# I.hat = lapply(R.JIVE.res$individual, svd, nu = r.I1, r.I2)
# 
# # R.JIVE Subject Scores
# R.JIVE.jnt.scores = svd(do.call(rbind, R.JIVE.res$joint), nu = r.J, nv = r.J)$v
# R.JIVE.ind.scores.X = I.hat[[1]][['v']]
# R.JIVE.ind.scores.Y = I.hat[[2]][['v']]
# 
# # R.JIVE loading estimates
# WJ = lapply(J.hat, function(x) x[['u']])
# WI = lapply(I.hat, function(x) x[['u']])
# 
# # Plots of R.JIVE estimates against true counterparts
# # layout(matrix(c(1,1,1:7), 3, byrow = TRUE), heights = c(.1, 0.4, 0.4))
# # plot(1, type = "n", axes=FALSE, xlab="", ylab=""); title(sub = "temp")
# layout(matrix(c(1:6,4,7,8),3, byrow = TRUE))
# plot(JntScores, R.JIVE.jnt.scores, xlab = "True Joint Scores",ylab = "R.JIVE Joint Scores", 
#      col = c(rep("orange",n), rep("green",n), rep("purple",n)),  
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntScores, R.JIVE.jnt.scores), 3)))
# plot(IndivScore.X, R.JIVE.ind.scores.X, xlab = "True Indiv 1 Scores", ylab = "ProJIVE Indiv 1 Scores", 
#      col = c(rep("orange",n), rep("green",n)),
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.X, R.JIVE.ind.scores.X), 3)))
# plot(IndivScore.Y, R.JIVE.ind.scores.Y, xlab = "True Indiv 2 Scores", ylab = "ProJIVE Indiv 2 Scores", 
#      col = c(rep("orange",n), rep("green",n)),
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivScore.Y, R.JIVE.ind.scores.Y), 3)))
# # plot(1,lwd=0,axes=F,xlab="",ylab="", "n")
# plot.new(); title("R.JIVE Results with \n R_J1^2=0.05")
# legend("left", paste("Comp.", 1:3), pch = 1, col  = c("orange", "green", "purple"),bty = "n" )
# plot(JntLd.X, WJ[[1]], xlab = "True Joint Loadings X", ylab = "R.JIVE Joint Loadings X", 
#      col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.X, WJ[[1]]), 3)))
# plot(JntLd.Y, WJ[[2]], xlab = "True Joint Loadings Y", ylab = "R.JIVE Joint Loadings Y", 
#      col = c(rep("orange",p1), rep("green",p1), rep("purple",p1)),  
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(JntLd.Y, WJ[[2]]), 3)))
# plot(IndivLd.X, WI[[1]], xlab = "True Individual Loadings X", ylab = "R.JIVE Individual Loadings X", 
#      col = c(rep("orange",p1), rep("green",p2)), 
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.X, WI[[1]]), 3)))
# plot(IndivLd.Y, WI[[2]], xlab = "True Individual Loadings Y", ylab = "R.JIVE Individual Loadings Y", 
#      col = c(rep("orange",p1), rep("green",p2)), 
#      sub = paste0("Chordal Norm = ", round(chord.norm.diff(IndivLd.Y, WI[[2]]), 3)))
# layout(1)
```

