---
title: "Problematic Identifiability Example"
author: "Benjamin Risk"
date: "4/15/2022"
output: html_document
---


Here, we present an example that shows issues that arise when 
$W_{Jk}$ and $W_{Ik}$ are not linearly independent:

   i) the joint versus individual variation is not identifiable, 
   
   ii) the individual components are not identifiable,
   
   iii) the number of individual components is not identifiable. 

To examine the invariance of the likelihood, it suffices to consider the covariance matrices. Recall $Cov(x_{ik}) = W_{Jk}W_{Jk}^\top + W_{Ik}W_{Ik}^\top + \sigma_k^2 I$. We can ignored $\sigma_k^2$ in the account that follows. 

# Generate loadings
```{r}
WJ1 = diag(5)[,c(1,2)]
WI1 = diag(5)[,c(3,4)]
```


For dataset 2, we will create a non-zero subspace in the  $col(W_{J2}) \cap col(W_{I2})$:

```{r}
WJ2 = diag(5)[,c(1,2)]
WI2 = diag(5)[,c(1,3)]
```

Consider the setting described in theorem 1. Let $T_2 = (T_1^{-1})^\top$. In order for the individual components of the alternative decomposition to be well-defined, their covariance matrix $W_{Ik}W_{Ik}^\top + W_{Jk}(I - T_k T_k^\top)W_{Jk}^\top$ needs to be positive semidefinite. If $col(W_{J2}) \cap col(W_{I2}) \ne \{0\}$, the eigenvalues of $W_{J2}(I-T_2 T_2^\top)W_{J2}^\top$ can be negative if the eigenvalue(s) of the matched directions in $W_{I2}$ are larger than the corresponding eigenvalue(s) of $W_{J2}(I - T_2 T_2^\top) W_{J2}^\top$. 

Then such a setting in turn allows a non-negligible contribution of $W_{J1}$ to the new individual subspace $W_{I1}^*$ when $I - T_1 T_1^\top$ is positive semidefinite.
```{r}
T1 = diag(c(sqrt(0.5+0.1),1)) 
T2 = t(solve(T1))
```

Construct the new joint loadings matrices:

```{r}
WJ1star = WJ1%*%T1
WJ2star = WJ2%*%T2
```

Construct the new individual components covariance matrix:

```{r}
covWI1star = WI1%*%t(WI1) + WJ1%*%(diag(ncol(WJ1))-T1%*%t(T1))%*%t(WJ1)
eig.WI1star = eigen(covWI1star)
#note no negative eigenvalues:
eig.WI1star$values
(rI1.star = sum(eig.WI1star$values>1e-8))
WI1star = eig.WI1star$vectors[,1:rI1.star]%*%diag(sqrt(eig.WI1star$values)[1:rI1.star])
```

We now have THREE non-zero eigenvalues, i.e., the number of individual components is not identifiable. This occurs because the overlap in the column spaces of $W_{J2}$ and $W_{I2}$ allows $(I - T_2 T_2^\top)$ to have negative eigenvalues ($T_2 = (T_1^{-1})^\top$), and since $I - T_1 T_1^\top$ is positive definite, $W_{J1}(I - T_1 T_1^\top) W_{J1}^\top$ has a non-negligible contribution of $W_{J1}$. 

We can make the situation even worse by rotating these components:
```{r}
Omat = svd(matrix(rnorm(rI1.star^2),rI1.star))$u
WI1star= WI1star%*%Omat
WI1star
```

Next, define the new loadings of the individual components of dataset 2:
```{r}
covWI2star = WI2%*%t(WI2) + WJ2%*%(diag(ncol(WJ2))-T2%*%t(T2))%*%t(WJ2)
eig.WI2star = eigen(covWI2star)
# check all positive:
eig.WI2star$values
WI2star = eig.WI2star$vectors[,1:2]%*%diag(sqrt(eig.WI2star$values)[1:2])
```


Finally, check covariance matrices are equivalent, which implies the likelihoods are equivalent:

```{r}
cov1 = WJ1%*%t(WJ1)+WI1%*%t(WI1)
cov2 = WJ2%*%t(WJ2)+WI2%*%t(WI2)

cov1star = WJ1star%*%t(WJ1star)+WI1star%*%t(WI1star)
cov2star = WJ2star%*%t(WJ2star)+WI2star%*%t(WI2star)

norm(cov1-cov1star,type = 'F')
norm(cov2-cov2star,type = 'F')
```

This demonstrates a ProJIVE model in which the joint versus individual variation is not identifiable, nor are the individual components identifiable. 

# Identifiability when $W_{Jk}$ and $W_{Ik}$ are linearly independent

When $col(W_{Jk}) \cap col(W_{Ik}) = \{0\}$, any $T_1 \notin \mathcal{O}$ will result in a negative eigenvalue for the covariance of the individual components, which is not a valid decomposition. It turns out that assuming linear independence leads to identifiability up to orthogonal transformation. 

Here is an example to see how the negative eigenvalues can arise, which implies that the $T_1$ matrix does not satisfy the constraints $W_{Ik}W_{Ik}^\top + W_{Jk}(I - T_k T_k^\top)W_{Jk}^\top \succeq 0$ for  $T_2 = (T_1^{-1})^\top$. .
```{r}
WJ1 = matrix(c(1,1,1,0,0),ncol=1)
WI1 = matrix(c(1,2,0,0,0),ncol=1)

WJ2 = matrix(c(0,0,0,1,0),ncol=1)
WI2 = matrix(c(0,0,0,0,1),ncol=1)

T1 = 1.01
T2 = 1/T1

WJ1star = WJ1*T1
WJ2star = WJ2*T2

# individual components covariance matrix:
covWI1star = WI1%*%t(WI1) + WJ1%*%(diag(ncol(WJ1))-T1%*%t(T1))%*%t(WJ1)
# check that it is positive semidefinite:
eigen(covWI1star)$values
```
